---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load_data,echo=TRUE}
fname <- "activity.csv"
zipfile<- "repdata-data-activity.zip"
url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if (!file.exists(fname)) {
  if (!file.exists(zipfile)) {
    download.file(url,destfile=zipfile)
  }
  unzip(zipfile)
}
act<-read.csv("activity.csv") 
```



## What is mean total number of steps taken per day?
first get the summary of steps taken each day
```{r sum_by_date,echo=TRUE}
act_sum_by_date <- aggregate(steps~date,data=act,sum)  
```

```{r hist_steps_per_day,echo=TRUE}
hist(act_sum_by_date$steps, 
     breaks=12,
     xlab="Total steps taken per day",
     main = "Personal Activity Over 2 Months")
```

get the mean and median from the summary figures
```{r summary of steps by date,echo=TRUE}
summary(act_sum_by_date)
```
```{r report_mean_median, echo=FALSE}
cat("The Mean of total number of steps taken per day", as.character(as.integer(mean(act_sum_by_date$steps))))

cat("The Median of total number of steps taken per day", as.character(median(act_sum_by_date$steps)))
```

## What is the average daily activity pattern?
first get the average of steps per 5 minute interval
```{r sum_by_interval,echo=TRUE}
act_sum_by_int <- aggregate(steps~interval,data=act,mean)

```
```{r Plot_average_by_interval, echo=TRUE}
plot(act_sum_by_int$interval,act_sum_by_int$steps,type="l",
     xlab="Interval",ylab="Average number of steps", 
     main = "Personal Activity Over 2 Months", 
     xaxt="n",yaxt="n")
axis(1,at=seq(0,2500,100),las=2)
axis(2,las=1)
```
```{r get_max_step_interval, echo=FALSE}
cat("The maximum number of steps is in interval",act_sum_by_int[which.max(act_sum_by_int$steps),"interval"])
```

## Imputing missing values
Calculate the number of missing values in the activity dataset
```{r count_na,echo=TRUE}
nrow(act[is.na(act$steps),])
```

Create a new dataset with missing values set to the mean for that interval
```{r impute_missing, echo=TRUE}
act_complete <- act
for (i in 1:nrow(act_complete)) {
    if (is.na(act_complete[i,"steps"]))
      act_complete[i,"steps"]<-act_sum_by_int[act_sum_by_int$interval==act_complete[i,"interval"],"steps"]
}
```

Aggregate the number of steps taken each day for the dataset with no missing values
```{r aggregate new dataset by day}
act_complete_sum_by_date <- aggregate(steps~date,data=act_complete,sum)
```
```{r hist_complete,echo=TRUE}
hist(act_complete_sum_by_date$steps,breaks=10)

```
get the mean and median from the summary figures
```{r summary_by_date_complete,echo=TRUE}
summary(act_complete_sum_by_date)
```
```{r report_mean_median_complete, echo=FALSE}
cat("After imputing missing values:")
cat("The Mean of total number of steps taken per day", as.character(as.integer(mean(act_complete_sum_by_date$steps))))

cat("The Median of total number of steps taken per day", as.character(as.integer(median(act_complete_sum_by_date$steps))))

cat("There does not seem to be much impact from imputing missing values.")
```

## Are there differences in activity patterns between weekdays and weekends?
Classify data into weekday and weekend
```{r weekday_or_end, echo=TRUE}
act_complete$daytype <- as.factor(ifelse(weekdays(as.Date(act_complete$date)) %in% c("Sunday","Saturday"),"weekend","weekday"))
act_c_daytype<-aggregate(steps~daytype+interval,data=act_complete,mean)
```
```{r daytype_plot,echo=TRUE}
library(lattice)
xyplot(steps~interval | daytype, data=act_c_daytype,type="l",layout=c(1,2))
```


